On a general class of discretely entropy conservative discontinuous Galerkin methods

### 2017-07-25 docs/splitform/main.tex
updating paper

\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}
\newtheorem*{remark}{Remark}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
%\usepackage{thmtools}
%\declaretheorem[style=definition,qed=$\blacksquare$,numberwithin=chapter]{definition}

@@

\date{}
\author{Jesse Chan}
\title{On discrete entropy conservation for a more general class of discontinuous Galerkin methods}

@@

\begin{abstract}
High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.  These include discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy} based on mass lumping and nodal collocation.  In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based projections and global DG differentiation operators.  The construct of these methods is independent of the choice of basis and accuracy of quadrature rule used, and includes DG methods involving generalized SBP operators without boundary nodes and certain dense-norm SBP operators.  Numerical experiments confirm the stability and high order accuracy of the proposed methods for the Burgers, shallow water, and compressible Euler equations.  
\end{abstract}

@@

\note{High order methods are good boilerplate}

\note{High order methods are not stable for nonlinear problems.  Stability of discretizations is a hot topic: artificial viscosity, polynomial de-aliasing,  \cite{karniadakis2013spectral}}

\cite{tadmor1987numerical, tadmor2003entropy}

Hughes entropy variables \cite{hughes1986new}

\note{Specify that entropy stability has been shown for mass lumped collocation DG discretizations.  Requirements are a diagonal norm SBP operators (or diagonal mass matrix) and collocated differentiation. }  \cite{gassner2016split,gassner2016well,wintermeyer2017entropy}  
\note{The construction of such operators on non-tensor product elements (such as simplices or pyramids) is non-trivial.} \cite{hicken2016multidimensional, chen2017entropy}.  \note{Constructing such operators for non-polynomial spaces (such as splines) is also difficult within the current framework.}

While DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.  Furthermore, GLL quadratures and diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.  Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher, cohen2001higher, hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.  Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.

\section{Entropy conservative diagonal norm collocation DG methods}

\note{Specify what we mean by diagonal norm collocation DG method.}

@@

We consider systems of nonlinear conservation laws in $\mathbb{R}^d$ with $n$ variables
\begin{align}
\pd{\bm{u}}{t} + \sum_{k=1}^d \pd{\bm{f_k(\bm{u})}}{\bm{x}_k} &= 0, \qquad \bm{u}(\bm{x},t) = (u_1(\bm{x},t),\ldots,u_n(\bm{x},t)).
\label{eq:pde}
\end{align}
where the fluxes $f_k(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}(\bm{x},t)$.
The Jacobian matrix $\bm{A}_k(\bm{u})$ is defined entrywise as
\[
\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
\]
We are interested in systems for which there exists a convex entropy function $U(\bm{u})$ such that  
\begin{equation}
U''(\bm{u})\bm{A}_k(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}_k(\bm{u})}^T.
\label{eq:entropysym}
\end{equation}
For systems with convex entropy functions, one can define entropy variables $\bm{v} = U'(\bm{u})$.  The convexity of the $U(\bm{u})$ guarantees that the mapping between conservative and entropy variables is invertible.

It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F_k(\bm{u})$ such that
\[
\bm{v}^T \pd{\bm{f}_k}{\bm{u}} = \pd{F_k(\bm{u})}{\bm{u}}^T.
\]
The flux function and entropy flux function are further related through the so-called entropy potential $\psi_k$ 
\[
\psi_k(\bm{v}) = \bm{v}^T\bm{f}_k(\bm{u}(\bm{v})) - F_k(\bm{u}(\bm{v})), \qquad \psi_k'(\bm{v}) = \bm{f}_k(\bm{u}(\bm{v})).
\]

When $\bm{u}$ is smooth, multiplying (\ref{eq:pde}) on the left by $\bm{v}^T = U'(\bm{u})^T$, applying the definition of the entropy flux and using the chain rule yields the conservation of entropy
\[
\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} = 0.
\]
More generally, it can be shown that physically relevant solutions of (\ref{eq:pde}) (defined as the limiting solution for an appropriately defined vanishing viscosity) satisfy the entropy inequality
\[
\pd{U(\bm{u})}{t} + \sum_{k=1}^d \pd{F_k(\bm{u})}{\bm{x}_k} \leq 0.  
\]
%Integration over $\mathbb{R}^d$ then shows that the entropy of the solution decreases in time.  
When $d = 1$, integrating the entropy inequality over $[-1,1]$ and using the definition of the entropy potential yields
\begin{equation}
\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}_k(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
\label{eq:consentropy}
\end{equation}

%\note{Wave equation example}

@@

Let $\widehat{x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.  Let $\ell_i(x)$ denote the nodal Lagrange basis defined at GLL points, and define the differentiation matrix $\bm{D}$ be defined as 
\[
\bm{D}_{ij} = \pd{\ell_j(x_i)}{x}.
\]
The matrix $\bm{D}$ maps from nodal values of a polynomial to nodal values of its derivative.  We also define the mass matrix $\bm{M}$
\[
\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x)\diff{x} \approx \delta_{ij} w_i,
\]
where the approximation results evaluating the integral using GLL quadrature.  For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ gives that $\bm{M}$ is a diagonal matrix.

@@

We derive a DG-SEM approximation by evaluating the above integrals using GLL quadrature.  We furthermore assume that the solution is approximated using a nodal basis at GLL points 
\[
\bm{u}(\bm{x},t) = \sum_{j=1}^{N+1} \bm{u}_j(t)\ell_j(x).  
\]
The resulting semi-discrete system yields a nodal collocation scheme for $\bm{u}_i$ 
\[
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}\otimes \bm{I}}_{ij}\bm{f}(\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))} = 0.
\]

This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).  Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.  
\begin{definition}
Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
\[
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
\]
The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$
\[
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
\]
Similarly, a flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is referred to as entropy stable if $\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) \leq (\psi_L - \psi_R)$.
\label{def:tadmor}
\end{definition}
An entropy conservative nodal collocation method can then be derived by modifying the flux derivative as follows
\begin{equation}
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
\label{eq:dgsem}
\end{equation}
The replacement of the derivative of the flux function with the flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.  We reproduce these proofs here, as we will refer to them in following sections.  
\begin{theorem}
The collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
\[
\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0,
\]
where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
\label{thm:dgsem}
\end{theorem}
\begin{proof}
Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
\[
\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
\]
where we have used continuity in time to arrive at the pointwise relation 
\[
\bm{v}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{\bm{u}}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{t}.
\]

The spatial derivative term can be manipulated using the summation by parts property $\bm{M}\bm{D} = \bm{S} = \bm{B}-\bm{S}^T$
\begin{align}
\sum_{i,j=1}^{N+1} \bm{v}_i^T 2 \bm{M}\bm{D}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{v}_i^T (\bm{S} + \bm{B} - \bm{S}^T)_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) \nonumber\\
&= \sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j). \label{eq:sbpentropy}
\end{align}
Rearranging indices and using symmetry of $\bm{f}_S(\bm{u}_i,\bm{u}_j)$ then yields that
\begin{align*}
\sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{v}_i - \bm{v}_j}^T\bm{f}_S(\bm{u}_i,\bm{u}_j)\\
&= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{\psi}_i-\bm{\psi}_j} = -\bm{1}^T \bm{S}\bm{\psi} = -\bm{1}^T\bm{B}\bm{\psi},
\end{align*}
where we have used that $\bm{S}\bm{1} = 0$ and the summation by parts property in the last line.

Using that $\bm{B}$ is diagonal, combining the latter term in (\ref{eq:sbpentropy}) with boundary terms involving the numerical flux yields
\[
\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
\]
where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
%Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
%\[
%\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0.
%\]
\end{proof}

\begin{remark}
The proof of entropy conservation can be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).  This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy,gassner2017br1,chen2017entropy}.  Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.  
\end{remark}

\section{Entropy conservative general DG methods}

>>>>

\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\theoremstyle{lemma}
\newtheorem{lemma}{Lemma}
\newtheorem*{remark}{Remark}
\theoremstyle{theorem}
\newtheorem{theorem}{Theorem}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}
%\usepackage{thmtools}
%\declaretheorem[style=definition,qed=$\blacksquare$,numberwithin=chapter]{definition}

@@

\date{}
\author{Jesse Chan}
\title{On a general class of discretely entropy conservative discontinuous Galerkin methods}

@@

\begin{abstract}
High order methods based on diagonal-norm summation by parts operators can be shown to semi-discretely conserve entropy for nonlinear systems of hyperbolic PDEs \cite{fisher2013high,carpenter2014entropy}.  These methods can be interpreted as mass lumped discontinuous Galerkin spectral element methods (DG-SEM) \cite{gassner2016split,gassner2016well,wintermeyer2017entropy, chen2017entropy}.  In this work, we describe how to extend discretely entropy conservative schemes for a more general class of DG methods using quadrature-based $L^2$ projections, flux differencing, and global DG differentiation operators.  This approach also recovers existing methods for Burgers' equation involving generalized SBP operators without boundary nodes and dense norm SBP operators.  Numerical experiments confirm the stability and high order accuracy of the proposed methods for the shallow water and compressible Euler equations.  
\end{abstract}

@@

Numerical simulations in engineering increasingly require higher accuracy without sacrificing computational efficiency.  High order methods are more accurate than low order methods per degree of freedom for sufficiently regular solutions, and provide one avenue towards improving fidelity in numerical simulations while maintaining reasonable computational costs.  High order methods which can accomodate unstructured meshes are particularly well-suited for problems with complex geometries.  Among such methods, high order discontinuous Galerkin (DG) methods are particularly well-suited to the solution of time-dependent hyperbolic problems on modern computing architectures \cite{hesthaven2007nodal, klockner2009nodal}.

The accuracy observed for high order methods can be attributed in part to their low numerical dissipation and dispersion properties \cite{ainsworth2004dispersive}.  While low numerical dissipation is advantageous for problems in linear wave propagation \cite{komatitsch1998spectral}, the high numerical dissipation present in low order schemes adds a stabilizing effect for nonlinear hyperbolic problems \cite{wang2013high}.  As a result, high order schemes for nonlinear conservation laws tend to require additional stabilization, including (but not limited to) filtering, slope limiting, artificial viscosity, and polynomial de-aliasing \cite{persson2006sub, krivodonova2007limiters, karniadakis2013spectral, hesthaven2007nodal, kirby2003aliasing}.  Moreover, stabilized numerical methods can still fail, requiring user intervention or heuristic modifications to achieve non-divergent solutions.

For linear wave propagation problems, numerical methods can be shown to be semi-discretely energy stable, even in the presence of curvilinear coordinates or variable coefficients \cite{warburton2013low, chan2016weight1, chan2016weight2, chan2017weight}.  This semi-discrete stability can be used to show that, under a stable timestep restriction (CFL), discrete solutions obtained by explicit time-stepping do not suffer from un-physical growth in time.  For nonlinear systems of conservation laws, typical methods for fluid dynamics do not admit formal proofs of stability.  However, progress has been made in designing provably discretely stable methods, beginning with the introduction of entropy conservative and entropy stable finite volume schemes by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.   These schemes rely on a specific entropy conservative flux which satisfies a condition involving the entropy variables and entropy potential, and have been extended to low order finite volume methods on unstructured grids \cite{ray2016entropy}.

The construction of entropy stable high order schemes was done much more recently for the compressible Euler and Navier-Stokes equations in \cite{fisher2013high, carpenter2014entropy} based on a spectral collocation approach on tensor product elements, which can be interpreted as a mass-lumped DG spectral element (DG-SEM) scheme.  The proof of entropy stability relies on the presence of a diagonal mass matrix, as well as the summation by parts property \cite{gassner2013skew}.  Similar entropy-stable schemes were constructed for the shallow water and MHD equations \cite{derigs2016novel, gassner2016well, wintermeyer2017entropy}.  More recently, entropy stable schemes were extended to unstructured triangular meshes in \cite{chen2017entropy} using the split-form methodology of \cite{gassner2016split} and multi-dimensional SBP operators \cite{hicken2016multidimensional}.

To the authors knowledge, the construction of entropy conservative and entropy stable schemes have relied on the use of lumped diagonal mass matrices and nodal collocation (diagonal norm SBP operators).  Entropy stable high order finite element and DG methods have been proposed \cite{hughes1986new}, but the proofs are often given at the continuous level, relying on exact integration or the chain rule, neither of which holds at the discrete level.  Thus, while DG-SEM based on flux differencing is entropy conservative, this is tied to a specific choice of basis and quadrature.  Diagonal-norm SBP operators can be constructed for triangles and tetrahedra \cite{chin1999higher, cohen2001higher, hicken2016multidimensional, chen2017entropy}; however, the number of nodal points for such operators is greater than the dimension of the natural polynomial approximation space.  Furthermore, appropriate point sets have only been constructed for $N \leq 4$ in three dimensions \cite{zhebel2014comparison}.  
%Furthermore, diagonal-norm SBP operators do not naturally generalize to triangles or tetrahedra \cite{helenbrook2009existence}.

This work focuses on the construction of entropy conservative high order DG schemes for systems of conservation laws.  In order to generalize beyond collocation DG and diagonal norm SBP operators, we will consider dense (non-lumped) mass matrices and quadrature rules of arbitrarily high accuracy (over-integration).  The outline of the paper is as follows: Section~\ref{sec:intro} will briefly review the construction of entropy conservative collocation methods on a single domain.  Section~\ref{sec:ecdg} will describe how to construct analogous entropy conservative methods on single domains.  Section~\ref{sec:ecdg2} will discuss extensions to multiple domains, including comparisons of different coupling terms and entropy stable fluxes.  We note that, for simplicity and clarity of presentation, the proposed methods and proofs are given in one dimension.  However, the extension to multiple dimensions (including simplicial elements) is straightforward and will be described in an upcoming manuscript.  
Finally, Section~\ref{sec:num} presents numerical results which verify the high order accuracy and discrete entropy conservation of the proposed methods.

%\note{The construction of such operators on non-tensor product elements (such as simplices or pyramids) is non-trivial.} \cite{hicken2016multidimensional, chen2017entropy}.  \note{Constructing such operators for non-polynomial spaces (such as splines) is also difficult within the current framework.}

\section{Entropy conservative diagonal norm collocation DG methods}
\label{sec:intro}

We will begin by reviewing continuous entropy theory, as well as existing high order methods which are provably entropy conservative at the discrete level.  These methods were introduced in \cite{fisher2013high, carpenter2014entropy} for tensor product elements, though we adopt methods of proof introduced more recently in \cite{gassner2017br1, chen2017entropy}.

%\note{Specify what we mean by diagonal norm collocation DG method.}

@@

%\note{Specify that all analysis is done in 1D but is extendable to higher dimensions!!  For simplicity and clarity of presentation, we will begin by focusing on the case $d = 1$.  }

We consider systems of nonlinear conservation laws in one dimension with $n$ variables
\begin{align}
%\pd{\bm{u}}{t} + \sum_{k=1}^d \pd{\bm{f_k(\bm{u})}}{x} &= 0, \qquad \bm{u}(x,t) = (u_1(x,t),\ldots,u_n(x,t)).
\pd{\bm{u}}{t} + \pd{\bm{f(\bm{u})}}{x} &= 0, \qquad \bm{u}(x,t) = (u_1(x,t),\ldots,u_n(x,t)).
\label{eq:pde}
\end{align}
where the fluxes $f(\bm{u})$ are smooth functions of the vector of conservative variables $\bm{u}({x},t)$.%$\bm{u}(x,t)$.
The Jacobian matrix $\bm{A}(\bm{u})$ is defined entrywise as
\[
%\LRp{\bm{A}_k(\bm{u})}_{ij} = \LRp{\pd{\bm{f}_k(\bm{u})}{\bm{u}_j}}_i.
\LRp{\bm{A}(\bm{u})}_{ij} = \LRp{\pd{\bm{f}(\bm{u})}{\bm{u}_j}}_i.
\]
We are interested in systems for which there exists a convex entropy function $U(\bm{u})$ such that  
\begin{equation}
%U''(\bm{u})\bm{A}_k(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}_k(\bm{u})}^T.
U''(\bm{u})\bm{A}(\bm{u}) = \LRp{U''(\bm{u}) \bm{A}(\bm{u})}^T.
\label{eq:entropysym}
\end{equation}
For systems with convex entropy functions, one can define entropy variables $\bm{v} = U'(\bm{u})$.  The convexity of the $U(\bm{u})$ guarantees that the mapping between conservative and entropy variables is invertible.

It can be shown (see, for example, \cite{mock1980systems}) that (\ref{eq:entropysym}) is equivalent to the existence of an entropy flux functions $F(\bm{u})$ such that
\[
\bm{v}^T \pd{\bm{f}}{\bm{u}} = \pd{F(\bm{u})}{\bm{u}}^T.
\]
The flux function and entropy flux function are further related through the so-called entropy potential $\psi_k$ 
\[
%\psi_k(\bm{v}) = \bm{v}^T\bm{f}_k(\bm{u}(\bm{v})) - F_k(\bm{u}(\bm{v})), \qquad \psi_k'(\bm{v}) = \bm{f}_k(\bm{u}(\bm{v})).
\psi(\bm{v}) = \bm{v}^T\bm{f}(\bm{u}(\bm{v})) - F(\bm{u}(\bm{v})), \qquad \psi'(\bm{v}) = \bm{f}(\bm{u}(\bm{v})).
\]

When $\bm{u}$ is smooth, multiplying (\ref{eq:pde}) on the left by $\bm{v}^T = U'(\bm{u})^T$, applying the definition of the entropy flux and using the chain rule yields the conservation of entropy
\[
\pd{U(\bm{u})}{t} + \pd{F(\bm{u})}{{x}} = 0.
\]
More generally, it can be shown that physically relevant solutions of (\ref{eq:pde}) (defined as the limiting solution for an appropriately defined vanishing viscosity) satisfy the entropy inequality
\[
\pd{U(\bm{u})}{t} + \pd{F(\bm{u})}{{x}} \leq 0.  
\]

We assume now that the domain is the interval $[-1,1]$.  Integrating the entropy inequality over this interval and using the definition of the entropy potential then yields
\begin{equation}
\int_{-1}^1 \pd{U(\bm{u})}{t} + \left.\LRp{\bm{v}^T\bm{f}(\bm{u}(\bm{v}))-\psi(\bm{v})}\right|_{-1}^1 \leq 0.  
\label{eq:consentropy}
\end{equation}
The focus of this work is the construction of methods which satisfy a discrete analogue of this entropy inequality.  We begin by reviewing existing discretely entropy conservative methods based on collocation and diagonal norm summation by parts (SBP) operators.

@@

Let ${x}_i \in [-1,1]$ denote the $(N+1)$ point Gauss-Lobatto-Legendre (GLL) quadrature rule with positive weights $w_1,\ldots,w_{N+1}$.  Let $\ell_i(x)$ denote the nodal Lagrange basis defined at GLL points, and define the differentiation matrix $\bm{D}$ be defined as 
\[
\bm{D}_{ij} = \pd{\ell_j(x_i)}{x}.
\]
The matrix $\bm{D}$ maps from nodal values of a polynomial to nodal values of its derivative.  We also define the mass matrix $\bm{M}$
\[
\bm{M}_{ij} = \int_{-1}^1 \ell_i(x)\ell_j(x)\diff{x} \approx \delta_{ij} w_i,
\]
where the approximation results from evaluating the integral using GLL quadrature.  For this choice of quadrature, the Kronecker property $\ell_j(x_i) = \delta_{ij}$ implies that $\bm{M}$ is a diagonal matrix whose entries are simply the quadrature weights.

@@

We derive a DG-SEM approximation by evaluating the above integrals using GLL quadrature.  We furthermore assume that the solution is approximated using a nodal basis at GLL points 
\[
\bm{u}(x,t) = \sum_{j=1}^{N+1} \bm{u}_j(t)\ell_j(x).  
\]
The resulting semi-discrete system yields a nodal collocation scheme for $\bm{u}_i$ 
\[
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{\bm{D}\otimes \bm{I}}_{ij}\bm{f}(\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))} = 0.
\]

This formulation conserves the quantities $\bm{u}_i$, but (except in specific cases) does not satisfy a discrete analogue of the statement of entropy conservation (\ref{eq:consentropy}).  Entropy conservative schemes may be constructed based on the SBP property and a ``flux differencing'' approach \cite{gassner2016split} involving a definition of entropy conservative/stable fluxes given by Tadmor \cite{tadmor1987numerical, tadmor2003entropy}.  
\begin{definition}
Let $\bm{f}_S(\bm{u}_L,\bm{u}_R)$ be a bivariate function such that it is symmetric and consistent with the flux function $\bm{f}(\bm{u})$
\[
\bm{f}_S(\bm{u}_L,\bm{u}_R) = \bm{f}_S(\bm{u}_R,\bm{u}_L), \qquad \bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})
\]
The numerical flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is entropy conservative if, for entropy variables $\bm{v}_L = \bm{v}(\bm{u}_L), \bm{v}_R = \bm{v}(\bm{u}_R)$
\[
\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) = (\psi_L - \psi_R), \qquad \psi_L = \psi(\bm{v}(\bm{u}_L)), \quad \psi_R = \psi(\bm{v}(\bm{u}_R)).  
\]
Similarly, a flux $\bm{f}_S(\bm{u}_L, \bm{u}_R)$ is referred to as entropy stable if $\LRp{\bm{v}_L - \bm{v}_R}^T \bm{f}_S(\bm{u}_L,\bm{u}_R) \leq (\psi_L - \psi_R)$.
\label{def:tadmor}
\end{definition}
An entropy conservative nodal collocation method can then be derived by modifying the flux derivative as follows
\begin{equation}
\pd{\bm{u}_i}{t} + \sum_{j=1}^{N+1}\LRp{\LRp{2\bm{D}\otimes \bm{I}}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) +\LRp{\bm{M}^{-1}\bm{B}}_{ij}(\bm{f}_i^* - \bm{f}(\bm{u}_i))}= 0.
\label{eq:dgsem}
\end{equation}
The replacement of the derivative of the flux function with flux differencing allows one to prove that (\ref{eq:dgsem}) conserves a discrete entropy \cite{gassner2017br1,chen2017entropy}.  We reproduce these proofs here, as we will refer to them in following sections.  
\begin{theorem}
The collocation method (\ref{eq:dgsem}) is discretely entropy conservative in the sense that
\[
\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0,
\]
where $\bm{1}$ denotes the vector of ones and $\bm{\psi}_i = \psi(\bm{v}_i)$ is the entropy potential in terms of the nodal value of $\bm{u}_i$.  
\label{thm:dgsem}
\end{theorem}
\begin{proof}
Let $\bm{v}_i = \bm{v}(\bm{u}_i)$ denote the entropy variables, defined in terms of the nodal values $\bm{u}_i$.  We multiply (\ref{eq:dgsem}) by $\LRp{\bm{M}\bm{v}}^T$.  
Using the fact that $\bm{M}$ is diagonal, the time derivative term can be manipulated to 
\[
\bm{v}^T\bm{M}\pd{\bm{u}}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \bm{v}_i^T \pd{\bm{u}_i}{t} = \sum_{i=1}^{N+1} \bm{M}_{ii} \pd{S_i}{t} = \bm{1}^T\bm{M}\pd{S_i}{t},
\]
where we have used continuity in time to arrive at the point-wise relation 
\[
\bm{v}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{\bm{u}}^T \pd{\bm{u}}{t} = \pd{U(\bm{u})}{t}.
\]

The spatial derivative term can be manipulated using the summation by parts property $\bm{M}\bm{D} = \bm{S} = \bm{B}-\bm{S}^T$
\begin{align}
\sum_{i,j=1}^{N+1} \bm{v}_i^T 2 \bm{M}\bm{D}_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{v}_i^T (\bm{S} + \bm{B} - \bm{S}^T)_{ij}\bm{f}_S(\bm{u}_i,\bm{u}_j) \nonumber\\
&= \sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j). \label{eq:sbpentropy}
\end{align}
Rearranging indices and using symmetry of $\bm{f}_S(\bm{u}_i,\bm{u}_j)$ then yields that
\begin{align*}
\sum_{i,j=1}^{N+1} \bm{S}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) - \bm{S}_{ji}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) &= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{v}_i - \bm{v}_j}^T\bm{f}_S(\bm{u}_i,\bm{u}_j)\\
&= \sum_{i,j=1}^{N+1} \bm{S}_{ij} \LRp{\bm{\psi}_i-\bm{\psi}_j} = -\bm{1}^T \bm{S}\bm{\psi} = -\bm{1}^T\bm{B}\bm{\psi},
\end{align*}
where we have used $\bm{S}\bm{1} = 0$ and the summation by parts property in the last line.

Using that $\bm{B}$ is diagonal, combining the latter term in (\ref{eq:sbpentropy}) with boundary terms involving the numerical flux yields
\[
\sum_{i,j = 1}^{N+1} \bm{B}_{ij}\bm{v}_i^T \bm{f}_S(\bm{u}_i,\bm{u}_j) + \bm{v}_i^T\bm{B}_{ij}\LRp{\bm{f}_i^* - \bm{f}(\bm{u}_i)} = \bm{1}^T\bm{B} \LRp{\bm{v}^T \bm{f}^*},
\]
where we have used the consistency of $\bm{f}_S(\bm{u},\bm{u}) = \bm{f}(\bm{u})$.  Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}).   
%Combining everything results in a discrete statement of the conservation of entropy (\ref{eq:consentropy}) 
%\[
%\bm{1}^T\LRp{\bm{M}\pd{U(\bm{u})}{t} + \bm{B}\LRp{\bm{v}^T\bm{f}^* - \bm{\psi}}} = 0.
%\]
\end{proof}
We note that, unlike the proof of continuous entropy conservation, the proof of discrete entropy conservation does not require the use of the chain rule, which may not hold at the discrete level.

The proof of entropy conservation can also be extended beyond a single domain by specifying appropriate simultaneous approximation terms (SBP-SAT).  This can be done by taking $\bm{f}_i^*$ to be the entropy conservative flux $f_S(u_L,u_R)$ at the interface between two elements \cite{carpenter2014entropy, gassner2017br1, chen2017entropy}.  Similarly, employing an entropy stable flux at element interfaces results in an entropy stable method which satisfies a global entropy inequality.  
%\end{remark}

\section{Entropy conservative general DG methods on a single domain}
\label{sec:ecdg}
